---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r}
require(BSDA)
library(BSDA)
require(EnvStats)   
library(EnvStats)
```

Ivan Nikolaichenko, Dmytro Hrebeniuk, Oleh Hykaviy

### Generating the data

```{r}

n <- 6+6+9


generate.a <- function(k){
  return(k*log((k^2)*n+pi) - as.integer(k*log((k^2)*n+pi)))
}

k = seq(100)
l = seq(101, 150)

k.data = sapply(k, generate.a)
l.data = sapply(l, generate.a)



x.k = qnorm(k.data)
y.k = qnorm(l.data)
```

## Problem1

$$
H_0\,: \mu_1 = \mu_2 \quad \textrm{vs.} \quad H_1\,:  \mu_1 \neq \mu_2 \,; \\
\quad \sigma^2_1 = \sigma^2_2 = 1
$$

1)  We use two sided $z\mbox{ - test}$ , because the is $\sigma^2$ known and $\mu$ unknown

2)  We know that $z(\mathbf x)$ is: $$
    Z = Z(\mathbf X,\mathbf Y) := \sqrt{\frac{mn}{(m+n)}} \frac{\overline{\mathbf x} - \overline{\mathbf y}}{\sigma}
    $$

    The pi-value for this test is $p(\mathbf x,\mathbf y) = 2 \, \Phi(-|z(\mathbf x,\mathbf y)|)$

```{r}
sigma <- 1

# the builded test

z.test(x.k, y.k, sigma.x = sigma, sigma.y = sigma)

# find it by ourselfs

Z <- sqrt(100*50/(100+50))*(mean(x.k)-mean(y.k))/sigma
cat("The z(x, y) is:", Z, "\n")


pi.value <- 2*pnorm(-abs(Z))
cat("The pi-value is:", pi.value)


```

```{r}
cat("The 2.5%  quantile is", qnorm(0.025), "\n")
cat("The 97.5% quantile is ", qnorm(0.975), "\n")
cat("our z =", Z)
```

As we see, the z-value is between the percentiles, so we should not reject $H_0$ given $\alpha = 0.05$

## Problem2

$$
 H_0\,: \sigma^2_1 = \sigma^2_2 \quad \textrm{vs.} \quad H_1\,: \sigma^2_1 > \sigma^2_2\,; \qquad \\ \mu_1 \textrm{ and } \mu_2 \textrm{ are unknown }
$$

1.We are using the f - test here(which is called var.test in R), because our $\mu_1$ and $\mu_2$ are unknown and we need to test variances of two samples.

2.Under the null hypothesis, the test statistics $F(\mathbf X,\mathbf Y):= \frac{S_{\mathbf Y\mathbf Y}/(m-1)}{S_{\mathbf X\mathbf X}/(n-1)}$ has the Fischer distribution $\mathscr{F}_{m-1,n-1}.$ Obviously under ð»1, ð¹ would assume smaller values. Thus the rejection region can be $$
C_\alpha := \{ \mathbf x\in\mathbb R^n,\mathbf y\in\mathbb R^m \mid f(\mathbf x,\mathbf y)\le f_{\alpha} \}
$$ where $f_\alpha$ is the quantile of level $\alpha$ for the Fischer distribution $\mathscr{F}_{n-1,m-1}$. In our case alpha is equal to 0,05.

The $p$-value of the test is $F_{\mathscr{F}}(f(\mathbf x,\mathbf y))$

```{r}
f = var(y.k)/var(x.k)
cat("Our f is", f)
cat("\n")

df1 = 49
df2 = 99
# pf(f, df1, df2, lower.tail = FALSE)

var.test(y.k, x.k, alterantive="less")
```

As we can see, the $p$-value \> $\alpha$ (0,508 \> 0,05), so we should not reject the $H_0$

```{r}
cat("The 2.5%  quantile is", qf(0.025, df1, df2), "\n")
cat("The 97.5% quantile is ", qf(0.975, df1, df2), "\n")
cat("our f =", f)
```

As we see, the $f$ is between the percentiles, so we should not reject $H_0$ given $\alpha = 0.05$

## Problem 3

The Kolmogorov-Smirnov Test (K-S test) compares your data with a known distribution and lets you know if they have the same distribution. This K-S test is also convenient because it is based on the empirical distribution function.

#### The Kolmogorov-Smirnov hypothesis :

$H_0\,$: The data follow a specified distribution

$H_a\,$: The data do not follow the specified distribution

The Kolmogorov-Smirnov test statistic is defined as:

$$d := sup_{tâˆˆR}|\widehat{F_x}(t) - F_0(t)|$$

And the rejection region is defined as: $$C_\alpha := \{\textbf{x}\in \mathbb{R}^n |\ d \geq d^{(n)}_{1-\alpha}\}$$ where $\alpha = 0.05$

### a)

$\left \{ x_{k} \right \}_{k=1}^{100}$ are normally distributed (with parameters calculated from the sample)

```{r}
ks.test(x.k, "pnorm")
hist(x.k)
```

The p-value is much larger than our $\alpha = 0.05$ which is why we should not reject the null hypothesis and assume that our data is normally distributed.

### b)

$\{|x_k| \}^{100}_{k=1}$ are exponentially distributed with Î» = 1

```{r}
ks.test(abs(x.k), "pexp")
hist(abs(x.k))

lambda = 1
x <- rexp(100,rate = 1/mean(abs(x.k)))
pts <- seq(-1,max(x),by=0.01)
plot(ecdf(x),col="darkblue")
lines(pts,pexp(pts,rate = lambda),col="red")

```

Because p-value is pretty larger than our $\alpha = 0.05$, we can assume that our data is not significantly different from exponential distribution distributed with Î» = 1

### c)

$\{x_k \}^{100}_{k=1}$ and $\{y_k \}^{50}_{l=1}$ have the same distributions

```{r}
ks.test(x.k, y.k)
plot(ecdf(x.k), col="red")
lines(ecdf(y.k), col="green")

```

Because p-value is larger than our $\alpha = 0.05$, we can assume that our data follows y.k distribution
